Solar wind discontinuities in the outer heliosphere: spatial distribution between 1 and 5 AU

### 2024-12-29 article.qmd
docs: incorporate Anton's revision

---
title: Spatial evolution of solar wind discontinuities in the outer heliosphere
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
affiliations:
  - id: ucla
    name: University of California, Los Angeles
execute: 
  eval: false
code-links:
  - href: ../
    text: Website
  - repo
  - href: https://www.overleaf.com/project/65b0a9c80a3616adf6e599e8
    text: Overleaf
abstract: We present a study of the spatial evolution of solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. We identify and analyze the properties of the discontinuities at different radial distances from the Sun. By differentiating the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance), we find that (1) the normalized occurrence rate of interplanetary discontinuities (IDs) decreases with radial distance from the Sun, following a $1/r$ relationship; (2) The thickness of IDs increases with radial distance, but when normalized to the ion inertial length, the thickness decreases; (3) The current density of IDs decreases with radial distance, but when normalized to the Alfvén current, the current density of IDs increases; (4) The distribution of the normalized thickness and current density of IDs remain constant over time at 1 AU on a yearly scale.
---

@@

- **What I did:** I studied the solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. I identified and analyzed the properties of the discontinuities at different radial distances from the Sun. I differentiated the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance).

- **What I found:** The normalized occurrence rate of IDs drops with the radial distance from the Sun, following a $1/r$ law. The thickness of IDs increases with the radial distance from the Sun, but after normalization to the ion inertial length, the thickness of IDs decreases. The current intensity of IDs decreases with the radial distance from the Sun, but after normalization to the Alfven current, the current intensity of IDs increases.

- **What it means:** The results of this study provide a better understanding of the solar wind discontinuities in the outer heliosphere and their spatial evolution. This information is important for understanding the dynamics of the solar wind and testing the local generation mechanism of discontinuities.

# Introduction

TODO: annihilation mechanisms and related references
TODO: References to stabliity and time scales

Rapid variations in interplanetary magnetic fields, commonly recognized as solar wind magnetic discontinuities [@colburnDiscontinuitiesSolarWind1966], embody localized transient rotations or jumps of the magnetic field. They are microscale structures observed throughout the heliosphere, from inner heliosphere [@liuCategorizingMHDDiscontinuities2022] to the heliosheath [@burlagaCurrentSheetsHeliosheath2011]. Despite being discovered over half a century ago, the specific origins and circumstances of their formation remain subjects of ongoing debate. 
Historically, it has been believed that discontinuities originate inside 1 AU, probably near or at the Sun, and are then convected passively outward by the solar wind [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001; @borovskyFluxTubeTexture2008]. However, the good agreement observed in the waiting-time analysis between solar wind data and magnetohydrodynamic (MHD) turbulence simulations [@grecoStatisticalAnalysisDiscontinuities2009] indicates that a significant portion of these discontinuities could be associated with flux-tube boundaries. These structures are likely intermittent, arising spontaneously and forming rapidly and locally within MHD turbulence. 
The annihilation mechanisms for these discontinuities are also uncertain. These discontinuities may be unstable features and some could rapidly wane or decay [@neugebauerTangentialDiscontinuitiesSolar1986]. Conversely, they can be very stable and persist over time intervals comparable to the convection time or the solar rotation period.
Further complicating this picture is the presence of different types of discontinuities in interplanetary space—specifically tangential and rotational discontinuities, which are classically regarded as static solutions to the MHD equations. These distinct types may arise from different sources and may evolve in unique ways. Investigating discontinuities across large radial distances has the potential to address these fundamental questions regarding their origin and evolution.

Discontinuities are not only an interesting topic in themselves but are also considered crucial for understanding the dynamics of the solar wind and cosmic rays. Theoretical models [@lerchePropagationMagneticDisturbances1975; @medvedevDissipativeDynamicsCollisionless1997] and MHD simulations [@grecoIntermittentMHDStructures2008; @yangFormationRotationalDiscontinuities2015] suggest that the formation and destruction of discontinuities are closely related to the nonlinear dynamics of Alfvén waves and Alfvénic turbulence. These nonlinear processes can create significant disturbances to the otherwise adiabatic evolution of the solar wind flow. 
These theoretical predictions are supported by spacecraft observations, which have revealed that the solar wind magnetic field follows the Parker model of the heliospheric current sheet only on average. Localized transient currents, which can be significantly more intense than those predicted by the Parker model, are carried by various discontinuities observed as strong variations in magnetic field components [@colburnDiscontinuitiesSolarWind1966; @burlagaMicroscaleStructuresInterplanetary1968; @turnerOrientationsRotationalTangential1971]. These discontinuities contribute significantly to the magnetic fluctuation spectra [@borovskyContributionStrongDiscontinuities2010] and play an important role in energy dissipation processes such as plasma heating [@osmanIntermittencyLocalHeating2012] and particle acceleration [@wentzelMotionMagneticDiscontinuities1964; @dmitrukTestParticleEnergization2004; @tesseinAssociationSuprathermalParticles2013]. 
More specifically, because tangential discontinuities act as boundaries separating different types of plasma and often involve oppositely directed magnetic fields on either side, they are potential regions where magnetic reconnection or merging can occur. Additionally, rotational discontinuities, being nonlinear large-amplitude Alfvén waves, are expected to significantly contribute to the diffusion coefficient for cosmic rays. Finally, these discontinuities can produce geomagnetic effects [@tsurutaniReviewInterplanetaryDiscontinuities2011] and impact space weather due to abrupt changes in external (solar wind) conditions.

In this study, we present a statistical analysis of discontinuities observed from Earth’s orbit to Jupiter, utilizing data collected by the Juno spacecraft [@boltonJunoMission2017] during its cruise phases (2011-2016), with its orbits shown in panel (a) of @fig-overview. The primary goal of this paper is to investigate the radial dependence of the occurrence rate and other properties of discontinuities, in order to gain insights into their evolution and origin. Continuous measurements at 1 AU from missions such as Wind, ARTEMIS, and STEREO have been utilized to distinguish temporal effects from spatial variations by examining the characteristics of discontinuities at two radial distances simultaneously. This distinction is crucial because temporal variations could easily be misinterpreted as spatial variations [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. Due to the potential misidentification of the normal direction of the magnetic field when using the minimum variance analysis (MVA) method [@hausmanDeterminingNatureOrientation2004; @liuFailuresMinimumVariance2023], we do not differentiate between the types of discontinuities, whether they are tangential, rotational, or either [@neugebauerReexaminationRotationalTangential1984].

Previous studies on the radial dependencies of solar wind discontinuities have been conducted by several authors [e.g., @marianiVariationsOccurrenceRate1973; @tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001]. However, the data used in these studies were rarely obtained in conjunction with measurements at fixed radial distances, and the methods employed to identify discontinuities were not robust for situations involving weak magnetic fields and long durations encountered at large radial distances. Additionally, some measurements at different radial distances were significantly separated in time, with limited time ranges covered [@sodingRadialLatitudinalDependencies2001]. Consequently, there remains ambiguity in interpreting the results of these studies.

The paper is structured as follows. First, we outline the missions, instruments, and data utilized in this study. Next, we provide a brief description of the method used to identify discontinuities and the model employed to estimate the plasma state at the Juno location, necessitated by the absence of plasma data during its cruise phase. In the subsequent sections, we present the results concerning the spatial variation of discontinuities, focusing on their occurrence rate and characteristics, such as spatial scales and current density. Finally, we discuss the implications of our findings in the context of the origin and evolution of these discontinuities in the solar wind.

@@

The solar wind plasma state, as evidenced by the sunspot number (referenced in panel (c) of @fig-overview), plays a crucial role in understanding the dynamics of discontinuities. During the initial phase of the JUNO mission, the sunspot number peaked, indicating a period of heightened solar activity. However, by the end of the mission's cruise phase, there was a significant decline in solar activity. This variation underscores the importance of calibrating the discontinuity properties in relation to solar activity levels to account for temporal fluctuations.

@@

Further analysis of @fig-overview highlights the significance of the heliographic longitudinal difference between the Juno mission and Earth as well as the STEREO-A mission. Specifically, when there is a substantial longitudinal difference between Juno and Earth, the difference between Juno and STEREO-A tends to be minimal, and vice versa. This longitudinal discrepancy between Earth and STEREO-A ensures comprehensive coverage of the plasma en route to Juno.

## Dataset

In this study, we analyze Juno observations during its cruise phase from August 25, 2011, to July 30, 2016, spanning approximately 1,800 days. Data collected from missions such as STEREO, ARTEMIS, and WIND complement Juno’s observations, providing a broader perspective on heliospheric conditions and enabling a multi-point analysis of discontinuity properties [@velliUnderstandingOriginsHeliosphere2020]. Specifically, we utilize magnetic fields from Juno [@boltonJunoMission2017], measured by the Juno Fluxgate Magnetometer (MAG) [@connerneyJunoMagneticField2017], alongside ion bulk velocity and plasma density with hourly resolution from a solar wind model at Juno’s location (described below). For ARTEMIS [@angelopoulosARTEMISMission2011], we use magnetic field data from the Fluxgate Magnetometer [@austerTHEMISFluxgateMagnetometer2008] and ion bulk velocity and plasma density derived from the Electrostatic Analyzer [@mcfaddenTHEMISESAPlasma2009]. For STEREO[@kaiserSTEREOMissionIntroduction2008], we incorporate magnetic fields from the magnetic field experiment on IMPACT [@acuñaSTEREOIMPACTMagnetic2008; @luhmannSTEREOIMPACTInvestigation2008] and ion bulk velocity and plasma density from PLASTIC [@galvinPlasmaSuprathermalIon2008]. For WIND, we include magnetic fields from the Magnetic Field Investigation (MFI) [@leppingWINDMagneticField1995] and ion bulk velocity and plasma density from the Solar Wind Experiment (SWE) [@ogilvieSWEComprehensivePlasma1995]. The table below summarizes the time resolution of magnetic field and plasma data for each mission. These synergistic observations are expected to enhance our understanding of solar wind discontinuities, including their radial distribution and evolution.

@@

However, since Juno does not provide direct plasma measurements, we will estimate the spatial scale (thickness) of discontinuities using solar wind plasma state data derived from solar wind propagation models. Specifically, we will employ the Two-Dimensional Outer Heliosphere Solar Wind Modeling (MSWIM2D) [@keeblerMSWIM2DTwodimensionalOuter2022] to determine the ion bulk velocity ($v$) and plasma density ($n$) at Juno’s location. This model, which utilizes the BATSRUS MHD solver[@tothAdaptiveNumericalAlgorithms2012], simulates the propagation of solar wind from 1 to 75 astronomical units (AU) in the ecliptic plane, effectively covering the region pertinent to our study. The MSWIM2D model provides output data with an hourly time resolution as shown in @fig-model. The comparison of magnetic field magnitudes from MSWIM2D with those measured by Juno, after averaging to the same time resolution, reveals a strong correlation, confirming the model’s reliability.

@@

## Methods

We employ Liu’s method [@liuMagneticDiscontinuitiesSolar2022] to identify discontinuities in the solar wind. This approach provides better compatibility for discontinuities with small magnetic field changes and demonstrates greater robustness for conditions in the outer heliosphere compared to traditional methods, namely the B-criterion [@burlagaDirectionalDiscontinuitiesInterplanetary1969] and the TS-criterion [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. For each sampling time $t$, we define three intervals: the pre-interval $[-1,-1/2]\cdot T+t$, the middle interval $[-1/2,1/2]\cdot T+t$, and the post-interval $[1/2,1]\cdot T+t$, where $T$ represents time lags. The magnetic field timeseries in these three intervals are labeled as ${\mathbf B}_-$, ${\mathbf B}_0$, and ${\mathbf B}_+$, respectively. To identify a discontinuity, the following three conditions must be met:

$$
\begin{aligned}
\sigma({\mathbf B}_0) &> 2\max\left(\sigma({\mathbf B}_-), \sigma({\mathbf B}_+)\right) \\
\sigma\left({\mathbf B}_-+{\mathbf B}_+\right) &>\sigma({\mathbf B}_-)+\sigma({\mathbf B}_+) \\
|\Delta {\mathbf B}| &>|{\mathbf B}_{bg}|/10
\end{aligned}
$$

Here, $\sigma$ denotes the standard deviation, ${\mathbf B}_{bg}$ represents the background magnetic field magnitude, and $\Delta {\mathbf B}={\mathbf B}(t+T/2)-{\mathbf B}(t-T/2)$. The first two conditions ensure that the field changes are distinguishable from stochastic fluctuations, while the third condition serves as a supplementary measure to reduce recognition uncertainty.

Notably, this method shares certain similarities with the Partial Variance of Increments (PVI) approach for identifying coherent structures, as both calculate indices based on standard deviations (with this method focusing on the standard deviation of the magnetic field, and PVI on magnetic field increments). However, this method is better suited for discontinuity identification by comparing data within neighboring time windows rather than across a correlation scale (an unknown quantity that may vary with radial distance and solar wind conditions). As a result, this approach effectively excludes wave-like coherent structures, eliminating the need for manual inspection and bypassing the necessity of grouping continuous clusters of high-PVI points as in PVI analysis.

Because the discontinuities may have different spatial scale at different radial distances from the Sun, we repeat the search procedure with time lags $T = 10, 20, 30, 40, 50, 60$ s, and combine the identified ID events together to establish the event list.
Discontinuities with spatial scales larger than $60$ s are not considered in this study, however it could be argued that they are not significant in the outer heliosphere if they follow the trend with decreasing number of IDs with increasing spatial scale. @fig-junoDistribution also indicates that the spatial scale of IDs does not vary significantly with the radial distance from the Sun.

@@

More than ? IDs are identified in the solar wind data from Juno, with ? IDs identified in 5 AU. Then for each discontinuity identified, we calculate the distance matrix of the time series sequence (distance between each pair of magnetic field vectors) to determine the leading edge and trailing edge of the discontinuity. After that, we use the minimum or maximum variance analysis (MVA) analysis [@sonnerupMinimumMaximumVariance1998; @sonnerupMagnetopauseStructureAttitude1967] to transform the magnetic field into the boundary normal (LMN) coordinate system. The main (most varying) magnetic field component, $B_l$, is then fitted by a hyperbolic tangent function to extract the parameters to properly describe the discontinuity @eq-fit. Then we combine the magnetic field data and plasma data to obtain the thickness and the current density of the discontinuity.

$$
B_l(t) = B_{l0} \tanh\left(\frac{t - t_0}{\Delta t}\right) + c_l
$${#eq-fit}

where $B_{l0}$ is the amplitude of the magnetic field change in the maximum variance direction, $t_0$ is the time of discontinuity, $\Delta t$ is the transition time, and $c_l$ is the offset of the magnetic field change (in the case of an ideal discontinuity with perfect rotational symmetry, $c_l$ should theoretically be zero. However, in the context of data processing, incorporating $c_l$ can enhance the fitting accuracy).

Previous studies have indicated that the minimum variance direction obtained through Minimum Variance Analysis (MVA) is frequently unreliable. Improved accuracy can be achieved by either using the cross product of the magnetic field measured at the boundaries as the normal direction [@knetterFourpointDiscontinuityObservations2004] or by enforcing conditions such as $Δ|B|/|B| > 0.05$ or $ω > 60°$ [@liuFailuresMinimumVariance2023]. The former approach requires a small normal component in the magnetic field, while the latter approach limits the dataset to a subset that satisfies these conditions. In our analysis, we applied both methods and observed no significant differences in the results.

@@

# Results

@@

Distribution of various properties of IDs observed by Juno grouped by the radial distance from the Sun. Panel (a) thickness, (b) normalized thickness, (c) current density, (d) normalized current density.
:::

@@

## Dependences of occurrence rate

>>>>

---
title: "Solar wind discontinuities in the outer heliosphere: spatial distribution between 1 and 5 AU"
author:
  - name: Zijin Zhang
    email: zijin@ucla.edu
    orcid: 0000-0002-9968-067X
    affiliation: University of California, Los Angeles
    roles: writing
    corresponding: true
  - name: Anton V. Artemyev
    email: aartemyev@igpp.ucla.edu
    orcid: 0000-0001-8823-4474
    affiliation: University of California, Los Angeles
  - name: Vassilis Angelopoulos
    orcid: 0000-0001-7024-1561
    affiliation: University of California, Los Angeles
  - name: Ivan Vasko
    affiliation: University of Texas at Dallas
affiliations:
  - id: ucla
    name: University of California, Los Angeles
execute: 
  eval: false
code-links:
  - href: ../
    text: Website
  - repo
  - href: https://www.overleaf.com/project/65b0a9c80a3616adf6e599e8
    text: Overleaf
abstract: Solar wind discontinuities (SWDs; also known as current sheets) are kinetic plasma structures contributing significantly to the spectrum of magnetic field fluctuations and to solar wind heating. Properties of SWDs varies significantly with the radial distance away from the Sun, and investigation of this radial evolution is important for revealing of SWD origin and their contribution to solar wind plasma processes. In this study we utilize quite unique dataset of Juno spacecraft observations of SWDs during cruise phase, when Juno was traveling from Earth ($\sim 1$AU) to Jupiter ($\sim 5$AU). We combine Juno measurements with Wind, ARTEMIS, and STEREO measurements of SWDs around $\sim 1$AU, and this combination allows us to separate temporal (with solar activity) and spatial (with radial distance) variations of SWD properties. We find that (1) the normalized to $1$AU occurrence rate of SWDs decreases with radial distance from the Sun, following a $1/r$ relationship; (2) the SWD thickness increases with radial distance, but when normalized to the ion inertial length, the thickness remains almost unchanged within $[1,5]$AU; (3) the SWD current density decreases with radial distance, but when normalized to the Alfvén current, the current density remains almost unchanged within $[1,5]$AU; (4) the distribution of SWD normalized thickness and current density remain constant over time (five year period of Juno cruise phase) at $1$ AU. We discuss these results in context of theories of SWD origin and dynamics.
# We present a study of the spatial evolution of solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. We identify and analyze the properties of the discontinuities at different radial distances from the Sun. By differentiating the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance), we find that (1) the normalized occurrence rate of interplanetary discontinuities (SWDs) decreases with radial distance from the Sun, following a $1/r$ relationship; (2) The thickness of SWDs increases with radial distance, but when normalized to the ion inertial length, the thickness decreases; (3) The current density of SWDs decreases with radial distance, but when normalized to the Alfvén current, the current density of SWDs increases; (4) The distribution of the normalized thickness and current density of SWDs remain constant over time at 1 AU on a yearly scale.
---

@@

<!-- - **What I did:** I studied the solar wind discontinuities in the outer heliosphere using data from the Juno spacecraft during its cruise phase. I identified and analyzed the properties of the discontinuities at different radial distances from the Sun. I differentiated the temporal effect (correlated with solar activity) and spatial variations (correlated with radial distance).
- **What I found:** The normalized occurrence rate of SWDs drops with the radial distance from the Sun, following a $1/r$ law. The thickness of SWDs increases with the radial distance from the Sun, but after normalization to the ion inertial length, the thickness of SWDs decreases. The current intensity of SWDs decreases with the radial distance from the Sun, but after normalization to the Alfven current, the current intensity of SWDs increases.
- **What it means:** The results of this study provide a better understanding of the solar wind discontinuities in the outer heliosphere and their spatial evolution. This information is important for understanding the dynamics of the solar wind and testing the local generation mechanism of discontinuities. -->
# Introduction

<!-- TODO: annihilation mechanisms and related references
TODO: References to stabliity and time scales -->

Rapid variations in the interplanetary magnetic field, commonly recognized as solar wind magnetic discontinuities (SWDs) [@colburnDiscontinuitiesSolarWind1966], embody localized transient rotations or jumps of the magnetic field. They are micro (ion kinetic) scale plasma structures observed throughout the heliosphere, from the inner heliosphere [@liuCategorizingMHDDiscontinuities2022] to the heliosheath [@burlagaCurrentSheetsHeliosheath2011]. Despite being discovered over half a century ago, the specific origins and circumstances of their formation remain subjects of ongoing debate \cite<see discussion in>[and references therein]{Vasko22:apjl,Vasko24:jgr}. 
Historically, SWDs have been believed to originate inside 1 AU, probably near or at the Sun, and are then passively convected outward by the solar wind [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001; @borovskyFluxTubeTexture2008]. However, the good agreement observed in the waiting-time analysis between solar wind data and magnetohydrodynamic (MHD) turbulence simulations [@grecoStatisticalAnalysisDiscontinuities2009] indicates that a significant portion of these discontinuities could be associated with flux-tube boundaries, intermittent structures, arising spontaneously and forming rapidly and locally within MHD turbulence. 
The annihilation mechanisms for these discontinuities are also uncertain. These discontinuities may be unstable features and some could rapidly wane or decay [@neugebauerTangentialDiscontinuitiesSolar1986]. Magnetic reconnection of SWDs is also considered a natural process that destroys them \cite{Gosling12,Wilson16:cs,else}. Conversely, SWDs can be very stable and persist over time intervals comparable to the convection time or the solar rotation period.
Further complicating this picture is the presence of different types of discontinuities in interplanetary space -— specifically tangential and rotational discontinuities \cite<see discussion in>{Neugebauer06,Artemyev19:jgr:solarwind,wangSolarWindCurrent2024}, which are classically regarded as static solutions to the MHD equations [@hudsonDiscontinuitiesAnisotropicPlasma1970]. These distinct types may arise from different sources and may evolve in unique ways. Investigating discontinuities across large radial distances has the potential to address these fundamental questions regarding their origin and evolution.

Discontinuities are not only interesting because they are nonlinear plasma structures widely observed in the heliosphere,  but are also considered crucial for understanding the dynamics of the solar wind plasma and cosmic rays. Theoretical models [@lerchePropagationMagneticDisturbances1975; @medvedevDissipativeDynamicsCollisionless1997] and MHD simulations [@grecoIntermittentMHDStructures2008; @yangFormationRotationalDiscontinuities2015] suggest that the formation and destruction of discontinuities are closely related to the nonlinear dynamics of Alfvén waves and Alfvénic turbulence. These nonlinear processes can create significant disturbances to the otherwise adiabatic evolution of the solar wind flow [@matteiniIonKineticsSolar2012; @tsurutaniReviewAlfvenicTurbulence2018].
These theoretical predictions are supported by spacecraft observations, which have revealed that the solar wind magnetic field follows the Parker model of the heliospheric current sheet only on average [@svirzhevskyHeliosphericMagneticField2021]. Localized transient currents carried by SWDs and significantly more intense than those predicted by the Parker model, are observed as strong variations in magnetic field components [@colburnDiscontinuitiesSolarWind1966; @burlagaMicroscaleStructuresInterplanetary1968; @turnerOrientationsRotationalTangential1971]. SWDs contribute significantly to magnetic fluctuation spectra [@borovskyContributionStrongDiscontinuities2010; @vaskoKineticScaleCurrentSheets2024] and play an important role in energy dissipation processes such as plasma heating [@osmanIntermittencyLocalHeating2012] and particle acceleration [@wentzelMotionMagneticDiscontinuities1964; @dmitrukTestParticleEnergization2004; @tesseinAssociationSuprathermalParticles2013]. 
More specifically, because tangential discontinuities act as boundaries separating different types of plasma and often involve oppositely directed magnetic fields on either side, they are potential regions where magnetic reconnection or merging can occur \cite<see reviews by>{Gosling12,Paschmann13}. Additionally, rotational discontinuities, being nonlinear large-amplitude Alfvén waves, are expected to contribute significantly to the scattering of energetic particles [@artemyevSuperfastIonScattering2020].
<!-- Finally, these discontinuities can produce geomagnetic effects [@tsurutaniReviewInterplanetaryDiscontinuities2011] and impact space weather due to abrupt changes in external (solar wind) conditions. -->

Investigation of the evolution of SWD properties with the radial distance (or alternatively, investigation of properties of SWD generation at different radial distances) has mostly been limited to comparison of SWD statistics collected by different missions during different intervals of solar activity [e.g., @marianiVariationsOccurrenceRate1973; @tsurutaniInterplanetaryDiscontinuitiesTemporal1979; @sodingRadialLatitudinalDependencies2001]. However, the data used in these studies were rarely obtained in conjunction with measurements at fixed radial distances, and the methods employed to identify discontinuities were not robust for situations involving weak magnetic fields and long durations encountered at large radial distances. Additionally, some measurements at different radial distances were significantly separated in time, with limited time ranges covered [@sodingRadialLatitudinalDependencies2001]. Consequently, there remains an ambiguity in interpreting the results of these studies.
In this study, we present a statistical analysis of discontinuities observed from Earth's orbit ($\sim 1$ AU) to Jupiter orbit ($\sim 5$ AU) by a single mission. We utilize data collected by the Juno spacecraft [@boltonJunoMission2017] during its cruise phases (2011-2016), with its orbits shown in panel (a) of @fig-overview. The primary goal of this study is to investigate the radial dependence of the SWD occurrence rate and properties, in order to gain insights into their evolution and origin. Continuous measurements of SWDs at 1 AU from missions such as Wind [@acunaGlobalGeospaceScience1995], ARTEMIS [@angelopoulosARTEMISMission2011], and STEREO [@kaiserSTEREOMissionIntroduction2008] have been utilized to distinguish temporal effects from spatial variations by examining the characteristics of discontinuities at two radial distances ($1$ AU and at Juno location) simultaneously. This distinction is crucial because temporal variations could easily be misinterpreted as spatial variations [see discussion in @tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. Due to the potential misidentification of the SWD normal plane with single spacecraft [see discussion in @hausmanDeterminingNatureOrientation2004; @liuFailuresMinimumVariance2023; @wangSolarWindCurrent2024], we do not differentiate between the types of discontinuities, whether they are tangential, rotational, or either [@neugebauerReexaminationRotationalTangential1984].

The paper is structured as follows. First, we outline the missions, instruments, and data utilized in this study (@sec-dataset). Next, we provide a brief description of the method used to identify discontinuities and the model employed to estimate the plasma state at the Juno location, necessitated by the absence of plasma data during its cruise phase (@sec-methods). In subsequent @sec-results, we present the results concerning the spatial variation of discontinuities, focusing on their occurrence rate and characteristics, such as spatial scales and current density. Finally, we discuss the implications of our findings in the context of the origin and evolution of these discontinuities in the solar wind (@sec-dependence).

@@

The solar wind plasma state, as evidenced by the sunspot number (referenced in panel (c) of @fig-overview), plays a crucial role in understanding the dynamics of discontinuities. During the initial phase (2011) of the Juno mission [@boltonJunoMission2017], the sunspot number peaked, indicating a period of heightened solar activity. However, by the end of the mission's cruise phase in 2016, there was a significant decline in solar activity. This variation underscores the importance of calibrating the discontinuity properties in relation to solar activity levels to account for temporal fluctuations.

@@

Further analysis of @fig-overview highlights the significance of the heliographic longitudinal difference between the Juno mission and Earth as well as the STEREO-A mission [@kaiserSTEREOMissionIntroduction2008]. Specifically, when there is a substantial longitudinal difference between Juno and Earth, the difference between Juno and STEREO-A tends to be minimal, and vice versa. This longitudinal discrepancy between Earth and STEREO-A ensures comprehensive coverage of the plasma en route to Juno.

## Dataset {#sec-dataset}

In this study, we analyze Juno observations during its cruise phase from August 25, 2011, to July 30, 2016, spanning approximately 1,800 days. Data collected from missions such as STEREO, ARTEMIS, and WIND complement Juno’s observations, providing a broader perspective on heliospheric conditions and enabling a multi-point analysis of discontinuity properties [@velliUnderstandingOriginsHeliosphere2020]. Specifically, we utilize magnetic fields from Juno [@boltonJunoMission2017], measured by the Juno Fluxgate Magnetometer (MAG) [@connerneyJunoMagneticField2017]. For ARTEMIS [@angelopoulosARTEMISMission2011], we use magnetic field data from the Fluxgate Magnetometer [@austerTHEMISFluxgateMagnetometer2008] and ion bulk velocity and plasma density derived from the Electrostatic Analyzer [@mcfaddenTHEMISESAPlasma2009]. For STEREO[@kaiserSTEREOMissionIntroduction2008], we incorporate magnetic fields from the magnetic field experiment on IMPACT [@acunaSTEREOIMPACTMagnetic2008; @luhmannSTEREOIMPACTInvestigation2008] and ion bulk velocity and plasma density from PLASTIC [@galvinPlasmaSuprathermalIon2008]. For WIND [@acunaGlobalGeospaceScience1995], we include magnetic fields from the Magnetic Field Investigation (MFI) [@leppingWINDMagneticField1995] and ion bulk velocity and plasma density from the Solar Wind Experiment (SWE) [@ogilvieSWEComprehensivePlasma1995]. The table below summarizes the time resolution of magnetic field and plasma data for each mission. These synergistic observations are expected to enhance our understanding of solar wind discontinuities, including their radial distribution and evolution.

@@

Since Juno does not provide direct plasma measurements during the cruise phase, we will estimate the spatial scale (thickness) of discontinuities using solar wind plasma state data derived from solar wind propagation models. Specifically, we will employ the Two-Dimensional Outer Heliosphere Solar Wind Modeling (MSWIM2D) [@keeblerMSWIM2DTwodimensionalOuter2022] to determine the ion bulk velocity ($v$) and plasma density ($n$) at Juno’s location. This model, which utilizes the BATSRUS MHD solver[@tothAdaptiveNumericalAlgorithms2012], simulates the propagation of solar wind from 1 to 75 astronomical units (AU) in the ecliptic plane, effectively covering the region pertinent to our study. The MSWIM2D model provides output data with an hourly time resolution as shown in @fig-model. The comparison of magnetic field magnitudes from MSWIM2D with those measured by Juno, after averaging to the same time resolution, reveals a strong correlation, confirming the model’s reliability.

@@

## Methods {#sec-methods}

We employ @liuMagneticDiscontinuitiesSolar2022 method to identify discontinuities in the solar wind. This method provides better compatibility for discontinuities with small magnetic field changes and demonstrates greater robustness for conditions in the outer heliosphere compared to traditional methods, namely the $B$-criterion [@burlagaDirectionalDiscontinuitiesInterplanetary1969] and the $TS$-criterion [@tsurutaniInterplanetaryDiscontinuitiesTemporal1979]. For each sampling time $t$, we define three intervals: the pre-interval $[-1,-1/2]\cdot T+t$, the middle interval $[-1/2,1/2]\cdot T+t$, and the post-interval $[1/2,1]\cdot T+t$, where $T$ represents time lags. The magnetic field time series in these three intervals are labeled as ${\mathbf B}_-$, ${\mathbf B}_0$, and ${\mathbf B}_+$, respectively. To identify a discontinuity, the following three conditions must be met:
$$
\begin{aligned}
\sigma({\mathbf B}_0) &> 2\max\left(\sigma({\mathbf B}_-), \sigma({\mathbf B}_+)\right) \\
\sigma\left({\mathbf B}_-+{\mathbf B}_+\right) &>\sigma({\mathbf B}_-)+\sigma({\mathbf B}_+) \\
|\Delta {\mathbf B}| &>|{\mathbf B}_{bg}|/10
\end{aligned}
$$
Here, $\sigma$ denotes the standard deviation, ${\mathbf B}_{bg}$ represents the background magnetic field magnitude, and $\Delta {\mathbf B}={\mathbf B}(t+T/2)-{\mathbf B}(t-T/2)$. The first two conditions ensure that the field changes are distinguishable from stochastic fluctuations, while the third condition serves as a supplementary measure to reduce recognition uncertainty [see discussion in @liuMagneticDiscontinuitiesSolar2022].

Notably, this method shares certain similarities with the Partial Variance of Increments (PVI) approach [@grecoPartialVarianceIncrements2017] for identifying coherent structures, as both calculate indices based on standard deviations (with this method focusing on the standard deviation of the magnetic field, and PVI on magnetic field increments). However, @liuMagneticDiscontinuitiesSolar2022 method is better suited for discontinuity identification by comparing data within neighboring time windows rather than across a correlation scale (an unknown quantity that may vary with radial distance and solar wind conditions). As a result, this approach effectively excludes wave-like coherent structures, eliminating the need for manual inspection and bypassing the necessity of grouping continuous clusters of high-PVI points as in PVI analysis.

Because the discontinuities may have different spatial scale at different radial distances from the Sun, we repeat the search procedure with time lags $T = 10, 20, 30, 40, 50, 60$ s, and combine the identified SWD events together to establish the event list.
Discontinuities with spatial scales larger than $60$ s are not considered in this study, however it could be argued that they are not significant in the outer heliosphere if they follow the trend with decreasing number of SWDs with increasing spatial scale [see @Vasko22:apjl; @Vasko24:jgr].
<!-- @fig-junoDistribution also indicates that the spatial scale of SWDs does not vary significantly with the radial distance from the Sun. -->

@@

More than ? SWDs are identified in the solar wind data from Juno, with ? SWDs identified within $[4, 5]$ AU. Then for each discontinuity identified, we calculate the distance matrix of the time series sequence (distance between each pair of magnetic field vectors) to determine the leading edge and trailing edge of the discontinuity. After that, we use the minimum variance analysis (MVA) analysis [@sonnerupMinimumMaximumVariance1998; @sonnerupMagnetopauseStructureAttitude1967] to transform the magnetic field into the boundary normal ($LMN$) coordinate system. The main (most varying) magnetic field component, $B_l$, is then fitted by a hyperbolic tangent function to extract the parameters to properly describe the discontinuity @eq-fit. Then we combine the magnetic field data and plasma data (from model) to obtain the thickness and the current density of the discontinuity.
$$
B_l(t) = B_{l0} \tanh\left(\frac{t - t_0}{\Delta t}\right) + c_l
$${#eq-fit}
where $B_{l0}$ is the amplitude of the magnetic field change in the maximum variance direction, $t_0$ is the time of discontinuity detection, $\Delta t$ is the transition time, and $c_l$ is the offset of the magnetic field change (in the case of an ideal discontinuity with perfect rotational symmetry, $c_l$ should theoretically be zero. However, in the context of data processing, incorporating $c_l$ can enhance the fitting accuracy).

Previous studies have indicated that the minimum variance direction obtained through MVA is frequently unreliable \cite<see discussion in>{knetterFourpointDiscontinuityObservations2004,wangSolarWindCurrent2024}. Improved accuracy can be achieved by either using the cross product of the magnetic field measured at the boundaries as the normal direction [@knetterFourpointDiscontinuityObservations2004] or by enforcing conditions such as $Δ|B|/|B| > 0.05$ or $ω > 60°$ [@liuFailuresMinimumVariance2023], where $\Delta |{\mathbf B}|$ is the change in the magnitude of the magnetic field and $\omega$ is the field rotation angle across the discontinuity. The former approach requires a small normal component (component along the normal to SWD plane) in the magnetic field, while the latter approach limits the dataset to a subset that satisfies these conditions. In our analysis, we applied both methods and observed no significant differences in the results.

@@

# Results {#sec-results}

@@

Distribution of various properties of SWDs observed by Juno grouped by the radial distance from the Sun. Panel (a) thickness, (b) normalized thickness, (c) current density, (d) normalized current density.
:::

@@

## Dependence of occurrence rate {#sec-dependence}
